PanGraph: scalable bacterial pan-genome graph construction

### 2023-03-03 reviewer_response.tex

\subsection*{Reviewer 3}

\reviewer{R3}{The article ``PanGraph: scalable bacterial pan-genome graph construction'' reports a new method for constructing a Pan-Genome Graph (PGG) in an efficient manner with an open and extractable grah structure. Unfortunately, the authors appear to either be unaware of or downplay the existence of previous methods for the same purpose.\textbf{ The authors need to survey the existing literature and compare and contrast with their method to show the utility of their approach.}\\
      For instance the authors state: ``Traditional computational approaches towards whole-genome analysis either scale poorly with the number of genomes, or treat genomes as dissociated ``bags of genes'', and thus are not suited for this new era.''. There are many pan-genome analysis tools which treat the pan-genome as a graph and generate a PGG. There are two primary differences in these existing tools: those which generate a graph at the nucleotide level with nodes representing basepairs (sometimes nonbranching runs of nodes are collapsed) and those where nodes represent larger features such as genes.\\
      PanGraph appears to wed these approaches to some extent in that the alignments are done at the nucleotide level and the features are extracted as a characteristic of the alignments.}

\responsedraft{
      We apologize if our previous discussion of the literature was incomplete, it was not our intention to downplay prior work.
      We thank the reviewer for highlighting additional important work and aspects that we previously didn't discuss.
      We have rewritten the part in the introduction to reflect these the work by these and other authors and added a few sentences to the discussion to further highlight previous and alternative approaches. We hope that these edits address the concerns of the reviewer.
      % SURVEY OF THE LITERATURE
      % PGG have been used to:
      % - improve annotations [sutton 2021] + [ggcaller]
      % MERTIS OF OUR METHOD
      % - de Bruijn graphs represent SNPs and large scale variation in the same way, as edges in the graph. We instead have this scale separation and encode nucleotide-level variation in alignments, and structural changes in the graph
      % - PGG based on genes are sensitive to annotation errors. We are based on homology.
      % - Scalable algorithm.
}

### 2023-03-03 main.tex
update

% TODO: SI Section IIA: superfluous?
% TODO: minimap2 asm10 superfluous?

@@

While easy to conceptualize, the construction of pangenome graphs has proven computationally challenging.
Colored generalizations of the de Bruijn graph-based assemblers have been successively used to build graphs from large sequence sets, although the underlying efficiency derives from a fixed k-mer size which prevents modelling long-range homology \cite{iqbal2012novo,muggli2017succinct}.
An orthogonal approach has been to formulate the inference of the pangenome graph as a multiple genome alignment.
However, current methods either scale poorly to large sets of genomes \cite{darling2010progressivemauve}, focus on comparisons across diverse sets of species across the tree of life at the cost of memory \cite{armstrong2020progressive}, or utilize a reference-guided approach by partitioning genomes first into annotated genes \cite{gautreau2020ppanggolin,colquhoun2021pandora}.

>>>>

@@

While easy to conceptualize, the construction of pangenome graphs has proven computationally challenging.
Colored and compacted generalizations of the de Bruijn graph-based assemblers have been successively used to build graphs from large sequence sets \cite{iqbal2012novo,muggli2017succinct}. Currently tools exist to build graphs containing thousands of isolates \cite{holley2020bifrost}, and have proven useful in problems such as the detection of core pangenome \cite{schulz2022sequence} or the improvement of gene prediction, annotation and clustering \cite{horsfield2023accurate}. However in these methods the graph structure encodes at the same time nucleotide-level variation and large-scale structural rearrangements in the genome.

An orthogonal approach has been to formulate the inference of the pangenome graph as a multiple genome alignment.
However, current methods either scale poorly to large sets of genomes \cite{darling2010progressivemauve}, focus on comparisons across diverse sets of species across the tree of life at the cost of memory \cite{armstrong2020progressive}, or utilize a reference-guided approach by partitioning genomes first into annotated genes \cite{gautreau2020ppanggolin,colquhoun2021pandora}.
